Supplementary Materials: Probing the Deuteron at Very Large Internal Momenta

### 2020-10-21 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/reply_to_RefereeA.txt

-------------
Referee A:
-------------
4) More general comment: Not enough information is given on the measurements. The data was taken using a new detection system
(SHMS at Hall C in 12 GeV setting) and the reader is referred only to unpublished thesis (with no Web link) for details.
In addition, the analysis itself is not transparent enough to the reader. More information should be provided in the form of
supplementary materials or similar. See details below.
----------
Response:
----------
I have published my thesis on arXiv and reference it in the PRL manuscript. I also added supplementary materials.

-------------
Referee A:
-------------
5) The scale of Fig. 2 is misleading as it is very hard to assess the quality of the agreement between the data and
theory in the 0 – 700 MeV/c region. Please add an insert with a blowup on that region.
----------
Response:
----------
I have added an inset with a blowup for Fig. 2, subplots (a) and (b).

-------------
Referee A:
-------------
6) I noticed (and appreciate) the availability of the crosse-section data in txt format and the same should be done
for the relevant calculations so that the interested reader can examine the results in more detail.
----------
Response:
----------
I have added additional raw data files with the theoretical calculations as part of the
supplemental material.

-------------
Referee A:
-------------
7) Page 1 introduction: Should place more emphasis on the importance to the broader physics community, as suggested above.
Elucidating the short-distance structure of nuclei and nuclear force via deuteron measurements is a fascinating problem of
a broad interest. However, the present work does not, in my opinion, meet this criterion.
----------
Response:
----------
As mentioned in response to (1), the introductory paragraph was re-written.

@@

-------------
Referee A:
-------------
8) Page 2 introduction "determining whether or to what extent the description of nuclei in terms of nucleon/meson degrees of
freedom is still valid before having to include explicit quark degrees of freedom”. Do the authors really think that (e,e’p)
measurement can address the question of nucleon/meson and partonic degrees of freedom? At the end of the day the data are compared
with calculations that have large freedom in their description of the off-shell electron-nucleon cross-section and even more so in
the modeling of the short-distance part of the NN interaction. Therefore, even if partonic physics becomes important at high-momenta,
it can probably be masked by effective nucleon- based interaction that will fit the data. In any case, it is ok if the authors want
to make such connections, however, it will probably be better to do so as part of a broader discussion of the relation between SRC
studies and the partonic structure of nuclei, e.g. by referencing the EMC effect and the EMC-SRC connections published in the last years.
----------
Response:
----------
This part the referee refers to was removed in the introductory paragraph.

-------------
Referee A:
-------------
9) The distributions of the fundamental measured quantities such as Q2, xB , ω, Pp should be shown. The measurements were done in 3 momentum bins.
The overlap and the matching between the bins should be shown. This can be a part of a supplementary materials document that allow the referees
and the reader to assess the quality of the data.
----------
Response:
----------
I have added the missing momentum distributions as well as the kinematic distributions for Q2, xB, w and Pp 
for each of the central settings (80, 580 and 750 MeV/c).  See Supplemental Material.

-------------
Referee A:
-------------
10) How the was the contribution of the target walls removed / subtracted?
----------
Response:
----------
The background from target cell wall was accounted for as an overall normalization systematic uncertainty.
(See Supplemental Material)

@@

-------------
Referee A:
-------------
11) “hits in three of the four scintillator planes.” At least three?
----------
Response:
----------
Yes, at least 3/4 planes. I have clarified this in the manuscript.

-------------
Referee A:
-------------
12) What about systematic uncertainties due to software cuts, acceptance corrections etc.?
----------
Response:
----------
A study of the systematic effects on the cross-section due to software cuts was carried out in Section 5.10 of Ref. https://arxiv.org/abs/2009.11343
The results showed that a reasonable variation of the software cuts had a negligible effect on the cross section.

An additional study on the spectrometers acceptance systematic error was carried out by M.K. Jones where systematic error of ~1.4 % was determined
and added quadratically to the final resut.

### 2020-10-23 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/reply_to_RefereeA_revised.txt

Light-front calculations are currently underway [M.Sargsian, Private communication, Oct. 2020) which we hope will be used 
to compare with our data for the large paper.

-------------
Referee A:
-------------
3) Nucleon-Nucleon interaction: Here the authors make the connection clear. However no modern EFT calculations are compared
with the new data. It is true that these are limited by strong regulators are high-momenta, however there are several interactions
such as the Norfolk potentials [M. Piarulli] that do suffer letter from such effects. In addition, even with high-momentum
regulators, it is interesting to observe out to what momenta can these interactions be coupled with a high- resolution one-body
interaction model and still explain experimental data. Recent analysis of nuclear form-factor data suggest this can be quite high.

In addition, there is no reference to relevant previous works that study nuclei heavier than deuterium. These include
A=3 [124 (2020) 212501; PRL 94 (2005) 082305; 94 (2005) 192302] and 12 [Nature 578 (2020) 540] that study the NN force
to short distance structure of SRC deuteron like pairs.
----------
Response:
----------
Similar to the above comment (2) regarding relativistic calculations to account for the effects of different
wave functions, it requires a complete calculation including the FSI effects, which is out of scope of the current paper. 
We hope that the theory groups who are collaborating with us will be able to provide such calculations which will be
presented in a larger paper. These studies are currently in progress.

-------------
Referee A:
-------------
4) More general comment: Not enough information is given on the measurements. The data was taken using a new detection system
(SHMS at Hall C in 12 GeV setting) and the reader is referred only to unpublished thesis (with no Web link) for details.
In addition, the analysis itself is not transparent enough to the reader. More information should be provided in the form of
supplementary materials or similar. See details below.
----------
Response:
----------
The authors' thesis has been published on the arXiv and a reference has been added to the PRL manuscript.
In addition supplementary materials have been added.

-------------
Referee A:
-------------
5) The scale of Fig. 2 is misleading as it is very hard to assess the quality of the agreement between the data and
theory in the 0 – 700 MeV/c region. Please add an insert with a blowup on that region.
----------
Response:
----------
An inset with a blowup for Fig. 2, subplots (a) and (b) has been added.

-------------
Referee A:
-------------
6) I noticed (and appreciate) the availability of the crosse-section data in txt format and the same should be done
for the relevant calculations so that the interested reader can examine the results in more detail.
----------
Response:
----------
Additional raw data files with the theoretical calculations as part of the supplemental material have been added.

@@

-------------
Referee A:
-------------
7) Page 1 introduction: Should place more emphasis on the importance to the broader physics community, as suggested above.
Elucidating the short-distance structure of nuclei and nuclear force via deuteron measurements is a fascinating problem of
a broad interest. However, the present work does not, in my opinion, meet this criterion.
----------
Response:
----------
As mentioned in response to (1), the introductory paragraph was re-written.

-------------
Referee A:
-------------
8) Page 2 introduction "determining whether or to what extent the description of nuclei in terms of nucleon/meson degrees of
freedom is still valid before having to include explicit quark degrees of freedom”. Do the authors really think that (e,e’p)
measurement can address the question of nucleon/meson and partonic degrees of freedom? At the end of the day the data are compared
with calculations that have large freedom in their description of the off-shell electron-nucleon cross-section and even more so in
the modeling of the short-distance part of the NN interaction. Therefore, even if partonic physics becomes important at high-momenta,
it can probably be masked by effective nucleon- based interaction that will fit the data. In any case, it is ok if the authors want
to make such connections, however, it will probably be better to do so as part of a broader discussion of the relation between SRC
studies and the partonic structure of nuclei, e.g. by referencing the EMC effect and the EMC-SRC connections published in the last years.
----------
Response:
----------
The authors believe that (e, e'p) data can contribute significantly to this difficult question, however, given the word limit of this paper, 
this section was removed in the introductory paragraph.

-------------
Referee A:
-------------
9) The distributions of the fundamental measured quantities such as Q2, xB , ω, Pp should be shown. The measurements were done in 3 momentum bins.
The overlap and the matching between the bins should be shown. This can be a part of a supplementary materials document that allow the referees
and the reader to assess the quality of the data.
----------
Response:
----------
The missing momentum distributions as well as the kinematic distributions for Q2, xB, w and Pp 
for each of the central settings (80, 580 and 750 MeV/c) have been added to the Supplemental Material.

@@

-------------
Referee A:
-------------
10) How the was the contribution of the target walls removed / subtracted?
----------
Response:
----------
The background from target cell wall was accounted for as an overall normalization systematic uncertainty.
(See Supplemental Material)

-------------
Referee A:
-------------
11) “hits in three of the four scintillator planes.” At least three?
----------
Response:
----------
Yes, at least 3/4 planes. This has been clarified this in the manuscript.

-------------
Referee A:
-------------
12) What about systematic uncertainties due to software cuts, acceptance corrections etc.?
----------
Response:
----------
A study of the systematic effects on the cross-section due to software cuts was carried out in Section 5.10 of Ref. https://arxiv.org/abs/2009.11343
The results showed that a reasonable variation of the software cuts had a negligible effect on the cross section.

An additional study on the spectrometers acceptance systematic error was carried out by M.K. Jones where a systematic error of ~1.4 % was determined
and added quadratically to the final result.

@@

-------------
Referee A:
-------------
13) What is the size of the radiation correction? Please detail on its calculation procedure.
----------
Response:
----------
See Supplemental Materials.

-------------
Referee A:
-------------
14) “a weighted average of the cross sections were taken in the overlapping regions of pr”. Show consistency. 
----------
Response:
----------
To clarify, a weighted average of the reduced cross sections were taken for overlapping regions in pr. See example
of overlapping pr bins in Fig. 9 of Supplemental Materials.

-------------
Referee A:
-------------
15) “cach” should be “each”.
----------
Response:
----------
This typo was not found in the manuscript.

@@

-------------
Referee A:
-------------
16) Fig. 2 and relevant discussion on Page 5: what about EFT calculations below and above their cutoffs?

Pages 5-6 include extensive phenomenological verbal description of Fig 2. However, what it misses a discussion
of the possible meaning of what is actually observed? For example: “the MS CD-Bonn model was able to partially
describe the data over a larger range...” – what does it mean? “all models were unable to describe the data”
above some momentum-what that means?
----------
Response:
----------
The comparison between the measured (reduced) cross sections and the currently available
complete and state of the art calculations illustrates the limit to which a non-relativistic
wave function that is the solution of the Schroedinger equation is valid and able to describe
experimental data that probe the high momentum region of the np system in the most direct way
possible.

As was mentioned before, the complete theoretical interpretation of these results is
out of the scope of this paper and is currently being investigated by the theory
groups we are collaborating with. Our goal was to present these new and
interesting results and make them accessible to the general public.

### 2020-10-23 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/arXiv_PRLsubmission_v2/supplemental.tex
created version 2 of arXiv PRL submittal package

\includegraphics[scale=0.24]{plots/Em_and_Q2_CUT_80MeV_35deg.png}

@@

\includegraphics[scale=0.24]{plots/deltaAcc_and_ZtarCUT_80MeV_35deg.png}

@@

\indent Figure \ref{fig:delta_Ztar_cuts} (left) shows a momentum acceptance ($\delta$) cut made on the HMS in the range $8\%\leq\delta\leq8\%$, where the optics reconstruction

@@

$|\delta|\lesssim$3$\%$ which is well within the reliable SHMS momentum acceptance range of $-10 \leq \delta \leq22 \%$. Figure \ref{fig:delta_Ztar_cuts} (right) shows reaction vertex cut

@@

\includegraphics[scale=0.23]{plots/coin_and_eCal_CUT_80MeV_35deg.png}

@@

coincidences is not observed. Figure \ref{fig:coin_ecal_cuts} (right) shows a particle identification cut on the SHMS calorimeter total energy

@@

\includegraphics[scale=0.24]{plots/collimator_CUT_80MeV_35deg_data.png}

@@

HMS can punch through the collimator with a probability that is related to the proton-Tungsten total cross section. Figure \ref{fig:data_coll_cuts} (right)

@@

\includegraphics[scale=0.24]{plots/collimator_CUT_80MeV_35deg_SIMC.png}
\caption{Same as Fig. \ref{fig:data_coll_cuts}, but for simulation.}

@@

\includegraphics[scale=0.17]{plots/d2_kin_thnq35.png}

@@

\includegraphics[scale=0.17]{plots/d2_kin_thnq45.png}

@@

\includegraphics[scale=0.33]{plots/tgt_bkg_d2_pm580_allthnq.png}

@@

  a 40 MeV range, each. (right) The SHMS $z$-reaction vertex corresponding to the specified region in the missing energy spectrum. Inset (left): Missing energy spectrum on a logarithmic
  scale.}

@@

\includegraphics[scale=0.3]{plots/Pr_and_Ps.png}

@@

\includegraphics[scale=0.27]{plots/RC_factor.png}

@@

\includegraphics[scale=0.27]{plots/BC_factor.png}

@@

%\indent A study of the sensitivity on the experimental cross section due to variations in the event selection cuts was carried out. 
%To determine if the variation in each of the cuts contributes to a systematic effect and whether this contribution is significant
%enough to be considered as a systematic error, we used the approach by R. Barlow described in Ref. \cite{barlow2002systematic}.\\
%\indent Consider a cross section measurement done two different ways (i.e., apply different cuts). Let the measurements and their
%statistical uncertainties be: ($\sigma^{\mathrm{exp}}_{\mathrm{1}}\pm\delta\sigma^{\mathrm{exp}}_{\mathrm{1}}$) and ($\sigma^{\mathrm{exp}}_{\mathrm{2}}\pm\delta\sigma^{\mathrm{exp}}_{\mathrm{2}}$)
%where one of the measurements is a subset of the other. The difference and its associated uncertainty can be expressed as,
%\begin{subequations}
%  \begin{align}
%    &\Delta \equiv \sigma^{\mathrm{exp}}_{\mathrm{1}} - \sigma^{\mathrm{exp}}_{\mathrm{2}}, \\
%    &\sigma^{2}_{\Delta} \equiv (\delta\sigma^{\mathrm{exp}}_{\mathrm{1}})^{2} - (\delta\sigma^{\mathrm{exp}}_{\mathrm{2}})^{2},
%  \end{align}
%\end{subequations}
%where the error of the difference between the two measurements is found by taking the difference of their variance. As demonstrated in Ref.\cite{barlow2002systematic}, this
%error accounts for the possible correlation between the two measurements. By taking the ratio
%\begin{equation}
%  R_{\mathrm{Barlow}} \equiv \frac{\Delta}{\sigma_{\Delta}},
%\end{equation}
%a criterion imposed on $R_{\mathrm{Barlow}}$ determines whether the difference is significant enough to be considered as a systematic error or sufficiently small that it may be
%ignored. This criterion requires knowledge of the correlation between the subsets, but in general, as suggested in Ref.\cite{barlow2017}: if $R_{\mathrm{Barlow}} < 2$ (or $\Delta <2\sigma_{\Delta}$)
%the test passes and if $R_{\mathrm{Barlow}} > 4$ (or $\Delta >4\sigma_{\Delta}$), the test fails and the discrepancy must be added as a systematic error. For $2<R_{\mathrm{Barlow}}<4$, a judgement must be made.\\
%\indent Figure \ref{fig:Em_sys} below shows an example of the systematic effects on the missing energy cut for $\theta_{nq}=35^{\circ}$ and $45^{\circ}$ over the full $p_{\mathrm{r}}$ range.
%In Fig. \ref{fig:Em_sys}, the different color groups represent the Barlow ratio evaluated at difference between the subset and full missing energy cut range. For mostly the entire momentum range,
%the Barlow ratio was kept within 2-4 standard deviations with the exception of a few outliers which can be understood from the fact that these might have very similar variances. These systematic
%studies were mainly done to check the stability of the event selection cuts and the effects of cut variation on the cross section were found to be negligible. Similar plots for the other event selection
%cuts systematics can be found in Section 5.10 of Ref. \cite{cyero_phdthesis}. 
%\clearpage
%\begin{figure}[!ht]
%\includegraphics[scale=0.4]{plots/Em_syst.png}
%\caption{Systematic effects of the missing energy cut on the data cross section for $\theta_{nq}=35^{\circ}$ (left) and $45^{\circ}$ (right). The inner (black dashed) and outer (black solid)
%  lines represent the $\Delta=\pm2\sigma_{\Delta}$ and $\pm4\sigma_{\Delta}$ boundaries, respectively.}
%\label{fig:Em_sys}
%\end{figure}
\indent The systematic effect on the cross sections due to model dependency of the radiative and bin-centering corrections was investigated (see Section 5.10 of Ref.\cite{cyero_phdthesis})

@@

\includegraphics[scale=0.37]{plots/rad_sys.png}

@@

\includegraphics[scale=0.37]{plots/bc_sys.png}

@@

effect in the measured cross sections and thus no systematic uncertainties due to software cuts were included in the final result. For a detailed discussion of these studies see Section 5.10 of Ref.\cite{cyero_phdthesis}.

>>>>

\includegraphics[scale=0.33]{Em_and_Q2_CUT_80MeV_35deg.png}

@@

\includegraphics[scale=0.33]{deltaAcc_and_ZtarCUT_80MeV_35deg.png}

@@

\indent Figure \ref{fig:delta_Ztar_cuts} (left) shows a momentum acceptance ($\delta$) cut made on the HMS in the range $-8\%\leq\delta\leq8\%$, where the optics reconstruction

@@

$|\delta|\lesssim$3$\%$ which is well within the reliable SHMS momentum acceptance range of $-10 \leq \delta \leq22 \%$. \\
\indent Figure \ref{fig:delta_Ztar_cuts} (right) shows a reaction vertex cut

@@

\includegraphics[scale=0.23]{coin_and_eCal_CUT_80MeV_35deg.png}

@@

coincidences is not observed. \\
\indent Figure \ref{fig:coin_ecal_cuts} (right) shows a particle identification cut on the SHMS calorimeter total energy

@@

\includegraphics[scale=0.24]{collimator_CUT_80MeV_35deg_data.png}

@@

HMS can punch through the collimator with a probability that is related to the proton-Tungsten total cross section.\\
\indent Figure \ref{fig:data_coll_cuts} (right)

@@

\includegraphics[scale=0.25]{collimator_CUT_80MeV_35deg_SIMC.png}
\caption{Same as Fig. \ref{fig:data_coll_cuts}, but for simulation.  The $z$-axis (color palette) is in arbitrary units.}

@@

\includegraphics[scale=0.17]{d2_kin_thnq35.png}

@@

\includegraphics[scale=0.17]{d2_kin_thnq45.png}

@@

\includegraphics[scale=0.33]{tgt_bkg_d2_pm580_allthnq.png}

@@

  a 60 MeV range, each.  Inset (left): Missing energy spectrum on a logarithmic scale. (right) The SHMS $z$-reaction vertex corresponding to
  the specified region in the missing energy spectrum.}

@@

\includegraphics[scale=0.3]{Pr_and_Ps.png}

@@

\includegraphics[scale=0.27]{RC_factor.png}

@@

\includegraphics[scale=0.27]{BC_factor.png}

@@

\indent The systematic effect on the cross sections due to model dependency of the radiative and bin-centering corrections was investigated (see Section 5.10 of Ref. \cite{cyero_phdthesis})

@@

\includegraphics[scale=0.37]{rad_sys.png}

@@

\includegraphics[scale=0.37]{bc_sys.png}

@@

effect in the measured cross sections and thus no systematic uncertainties due to software cuts were included in the final result. For a detailed discussion of these studies see Section 5.10 of Ref. \cite{cyero_phdthesis}.

### 2020-10-23 MY_PUBLICATIONS/PRL/FINAL_DRAFT/PRL_Submission/prl_response/prl_reply/supplemental/supp.tex
created version 2 of arXiv PRL submittal package

\includegraphics[scale=0.33]{plots/coin_and_eCal_CUT_80MeV_35deg.png}
\caption{Event selection cuts on the electron-proton ($ep$) coincidence time (left) and total deposited energy on calorimeter normalized by the particle track momentum (right).}

@@

at least 3 out of 4 scintillator planes between both spectrometers. The spectrum is very clean as the typical 2 ns beam bunch structure due to accidental
coincidences is not observed. The out-of-time events observed at the tails can originate from the radiation in the Hall partially entering the detector huts and forming
a trigger in coincidence with the other arm. Figure \ref{fig:coin_ecal_cuts} (right) shows a particle identification cut on the SHMS calorimeter total energy

@@

\includegraphics[scale=0.25]{plots/collimator_CUT_80MeV_35deg_data.png}

@@

  a 40 MeV range, each.  Inset (left): Missing energy spectrum on a logarithmic scale. (right) The SHMS $z$-reaction vertex corresponding to

>>>>

\includegraphics[scale=0.23]{plots/coin_and_eCal_CUT_80MeV_35deg.png}
\caption{Event selection cuts on the electron-proton ($ep$) coincidence time (left) and total deposited energy on calorimeted normalized by the particle track momentum (right).}

@@

at least 3 out of 4 scintillator planes between both spectrometers. The spectrum is very clean as the typical 4 ns beam bunch structure due to accidental
coincidences is not observed. \\
\indent Figure \ref{fig:coin_ecal_cuts} (right) shows a particle identification cut on the SHMS calorimeter total energy

@@

\includegraphics[scale=0.24]{plots/collimator_CUT_80MeV_35deg_data.png}

@@

  a 60 MeV range, each.  Inset (left): Missing energy spectrum on a logarithmic scale. (right) The SHMS $z$-reaction vertex corresponding to
